Motmot: A Distributed Chat Protocol

### 2012-05-04 paper/paper.tex
Discovery server

\Motmot is a new distributed text chat protocol based on Paxos. Unlike previous
chat protocols like XMPP, OSCAR, and IRC, \Motmot provides guaranteed message
delivery, total message ordering, end-to-end encryption, peer authentication,
and message deniability.
% TODO: write more

@@

>>>>

\Motmot is a new distributed text chat protocol based on Paxos. Unlike previous
chat protocols like XMPP, OSCAR, and IRC, \Motmot provides guaranteed message
delivery, total message ordering, end-to-end encryption, peer authentication,
and message deniability.
% TODO: server federates, etc.
% TODO: no xml! It's a feature
% TODO: write more

@@

The discovery server performs several tasks in the system: authentication and
identity federation; presence management; and connection brokering.

\subsection{Identity}

Much like other federated protocols like XMPP and email, identity handles are
based upon DNS. Every domain name with a SRV record can act as a \Motmot
identity realm, and give out email-like addresses (an \verb`addr-spec` as
defined in RFC 2822) for \Motmot communication.

To grant a user an identity on a realm, the user first presents a username and
password to the server, after which the server delegates authority for that name
to the client by signing that client's key. This key-signing process is the
method by which peers can authenticate securely to each other without a
centralized broker.

\subsection{Presence}

In chat protocols, presence is the system by which users can specify a list of
buddies to publish their online status to. The \Motmot discovery server provides
a presence lookup service, by recording the status of all of the clients
connected to it. Servers then push presence updates to all of the other servers
on any user's buddy list, which in turn push the notification to the relevant
clients.

This subsystem system does not need to be robust or reliable, since all pushed
state can be recreated from other sources, and since asynchronously pushed
presence notifications do not need to be perfectly reliable to ensure a usable
user interface.

\subsection{Brokering}

Finally, the server must perform message brokering. For clients behind firewalls
or routers which perform network address translation (NAT), direct connections
are not possible. In such a case, the discovery server might be forced to assist
in connecting the clients.

Technologies like STUN, TURN, and ICE provide implementations of ``NAT poking,''
and while mechanisms for interfacing with such servers were considered in the
design of the server, performing brokering itself was outside the scope of this
version of the protocol.

The \Motmot discovery server was implemented in Python using several open source
libraries. The server itself is a custom RPC-like reactor/dispatcher written
% TODO: cite gevent
using gevent, a coroutine-based networking library. For our communication
% TODO: cite msgpack
protocol, we use msgpack to serialize messages. This allows for easy
cross-language communication with clients and other servers, since
implementations are available for most popular languages.

We found that these two libraries provided good abstractions for the task of
implementing a server. Gevent's coroutines provided an extremely useful
concurrency model: they were executed in a single thread, so synchronization
around global objects was not necessary, while avoiding the complexity (and
unnecessary indentation) of a callback-based system. Msgpack also provided an
excellent abstraction, allowing the server to assume that arbitrary data
structures could be written to and read from the wire. The ad-hoc nature of the
format also helped speed development of the server and protocol.

% TODO: cite pyopenssl
Finally, we used openssl (and the pyOpenSSL library in particular) to perform
certificate operations

% TODO: EJ?
>>>>>>> Discovery server
